# Potts Model on a free-scale network

The aim of this task is to perform a simulation of the Potts model on a complex network, to examine its critical behavior, specifically how the order parameter varies in response to changes in the control parameter.

### Libraries

```{r}
library(igraph) |> suppressPackageStartupMessages()
library(ggplot2) |> suppressPackageStartupMessages()
library(RColorBrewer) |> suppressPackageStartupMessages()
library(rgl)|> suppressPackageStartupMessages()
```

### Generation of the free-scale network

##### WARNING:

1)  Often p indicates the number available states but in this simulation the number of states is: p+1.

2)  This code can endle at most six Potts states. To add more than 6 states the code must be modified.

```{r}
N <- 100 # number of nodes 
gamma <- 3.5 # exponent 
p <- 2 # nuber of states - 1 

# Power-law degree distribution
degs <- sample(2:100, N, replace=TRUE, prob=(2:100)^-gamma)
# correct the degree sequence if its sum is odd
if (sum(degs) %% 2 != 0) { degs[1] <- degs[1] + 1 }
random_states <- sample((0:p), N, replace = TRUE) 
# Creation of the graph 
h <- sample_degseq(degs, method="simple.no.multiple") %>%  
  set_vertex_attr("state", value = random_states)   # normal conf. model
```

### Plot of the graph

```{r}
plot(h, 
     vertex.label=NA, 
     vertex.color="green",
     vertex.size=2,
     edge.color="red",
     edge.width=1)
```

### Definition of some functions:

##### delta(): is a simple delta function

##### ene_iniz(): is function that computes the initial energy of the system

```{r}
#
delta <- function(a, b) {
  ifelse(a == b, return (1), return (0))
}
#
ene_iniz <- function(h) {
  energy <- 0
  for(i in seq(1: vcount(h))) {
    nn <- neighbors(h, i, mode = c( "all"))
    stateNode <- V(h)$state[i]
    for(k in seq(1:length(nn))) {
    energy<- energy - delta(stateNode, V(h)$state[nn[k]])
    }
  }
  return(energy)
}
```

### Simulated annealing:

```{r}

#MAIN

temperature <- 3
cicli <- 200 
energy <- ene_iniz(h)
# initialization 
n0 <- 0
n1 <- 0
n2 <- 0
n3 <- 0
n4 <- 0
n5 <- 0
counter <- 0
counter2 <- 0
magnetization <- 0
stimE <- 0
stimE2 <- 0
temp <- c()
magn <- c()
#
repeat{
  counter <- counter + 1
  for (passi in 1:cicli) {
    e1 <- 0
    e2 <- 0
    de <- 0
    # select a random node 
    random <- sample(1:vcount(h), 1)
    stateNode1 <- V(h)$state[random]
    nn <- neighbors(h, random, mode = c( "total"))
    # compute the energy of the node before the flip
    for(node in 1:length(nn)) {
      e1 <- e1 - delta(stateNode1, V(h)$state[nn[node]])
    }
    # flip
    randomState <- sample(0:p, 1)
    # the node must change state 
    while (randomState == stateNode1) {
        randomState <- sample(0:p, 1)
    }
    V(h)$state[random] <- randomState
    stateNode2 <- V(h)$state[random]
    # compute the energy of the node after the flip
    for(node in 1:length(nn)) {
      e2 <- e2 - delta(stateNode2, V(h)$state[nn[node]])
    }
    # evaluate the delta energy
    de <- e2 - e1
    energy <- energy + de
    # Metropolis step condition 
    if ((de) >= 0) 
    {
    stochastic <- runif(1)
    if (exp(-(de) / temperature) < stochastic) 
      {
        # if the step is not accepted, restore the previous situation 
        V(h)$state[random] <- stateNode1
        energy <- energy - de
      }
    }
  } # end of the cicle from 1 to cicli 
  temperature <- temperature - 0.05
  cat ("Temperatura", temperature, "Energia:", energy, "\n")
  # more steps at the temperature of interest 
  if((counter %% 10 == 1)) 
  {
    cat("****BREAK****", "\n")
    # termalization of the system 
    for(steps in 1:(10^4)) 
    {
      # repetition of the Metropolis step 
      e1 <- 0
      e2 <- 0
      de <- 0
      random <- sample(1:vcount(h), 1)
      stateNode1 <- V(h)$state[random]
      nn <- neighbors(h, random, mode = c( "total"))
      for(node in 1:length(nn)) {
        e1 <- e1 - delta(stateNode1, V(h)$state[nn[node]])
      }
      # flip
      randomState <- sample(0:p, 1)
      while (randomState == stateNode1) {
        randomState <- sample(0:p, 1)
      }
      V(h)$state[random] <- randomState
      stateNode2 <- V(h)$state[random]
      for(node in 1:length(nn)) {
        e2 <- e2 - delta(stateNode2, V(h)$state[nn[node]])
      }
      de <- e2 - e1
      energy <- energy + de
      # Metropolis step condition 
      if ((de) >= 0) 
      {
        stochastic <- runif(1)
        if (exp(-(de) / temperature) < stochastic) 
        {
         V(h)$state[random] <- stateNode1
         energy <- energy - de
        }
      }
      # computation of the magnetization 
      if (steps %% 1000 == 1) {
      # counting the state 
      mask_vector0 <- V(h)$state == 0
      mask_vector1 <- V(h)$state == 1
      mask_vector2 <- V(h)$state == 2
      mask_vector3 <- V(h)$state == 3
      mask_vector4 <- V(h)$state == 4
      mask_vector5 <- V(h)$state == 5
      #
      n0 <- sum(mask_vector0)
      n1 <- sum(mask_vector1)
      n2 <- sum(mask_vector2)
      n3 <- sum(mask_vector3)
      n4 <- sum(mask_vector4)
      n5 <- sum(mask_vector5)
      #
      maxn <- max(n0, n1, n2, n3, n4, n5)
      magnetization <- magnetization + (((p+1) / (vcount(h)) * maxn - 1)/(p))
      counter2 <- counter2 + 1
      #
      n0 <- 0
      n1 <- 0
      n2 <- 0
      n3 <- 0
      n4 <- 0
      n5 <- 0
    }  
  } # end of the loop over the iteration at the interested temperature 
  # save the temp & magnetization
  temp <- append(temp, temperature)
  magn <- append(magn, magnetization/counter2)
  # feedback for the user
  cat("Magnetization", magnetization/counter2) 
  magnetization <- 0
  counter2 <- 0
  stimE <- 0
  stimE2 <- 0
  }
if(temperature < 0.1) {break}
}
```
