## Song-Havlin-Makse self-similar model

The aim of this task is to implement the growth mechanism behind the fractal, or self - similar, networks and to study its fractal properties.

### Generation of the network

In order to practically implement the model introduced before, we follow this algorithm based on two growth processes called Mode I and Mode II (see theory). The growth algorithm swtich from Mode I and Mode II according to probability e. In the following we show this method lead us to a small - world fractal network.

##### Mode I with prob = e.

##### Mode II with prob = 1 - e.

```{r}
library(igraph) |> suppressPackageStartupMessages()
library(ggplot2) |> suppressPackageStartupMessages()
#
# Main to generate a network with a multiplicative process and preferential attachment
#
# Model I
N_it <- 2  # Number of iterations
m <- 2     # Multiplicative factor 
e <- 0.8   # probability 
nn_list <- list() 
# Initialization of the starting graph as a star 
g <- make_graph(~ 1--2, 1--3, 1--4, 1--5)
diameter <- list()
numberNodes <- list()
#
# loop over the number of iterations 
for(t in 1:N_it) { 
  V(g)$old <- TRUE
  V(g)$busy <- TRUE
  V(g)$done <- FALSE
  degree_sequence <- degree(g)
  for( s in (1:vcount(g) ) ) {
  nn_list[[length(nn_list) + 1]] <- as_ids(neighbors(g, s, mode = c( "all")))
  }
  # MODE I
  for (node in 1:vcount(g)) {
    vcount(g)
    node_degree <- degree(g, v = node)
    mk = m*node_degree
    for( j in seq(1:mk)){
      g <- add.vertices(g, nv = 1, old = FALSE, done = TRUE, busy = FALSE, name = vcount(g) + 1)
      new_node <- vcount(g)
      # connection with child nodes 
      g <- add_edges(g, edges = c(node, new_node ))
    }
  }
  # loop over all the nodes
  for(v1 in 1:vcount(g)) { 
    # CHECK -- If the node already exists in the previous iteration
    if(V(g)$old[v1] == TRUE) { 
      # loop over its old neighbors
      for(old_nei in nn_list[[v1]]) { 
        # CHECK -- If the old node's neighbors have not already been considered
        if (V(g)$done[as.integer(old_nei)] == FALSE) { 
          # introduction of stochasticity
          random <- runif(1)
          if (random > e) {
            # MODE II 
            del_edge <- get.edge.ids(g, c(as.integer(v1), as.integer(old_nei)))
            g <- delete_edges(g, as.integer(del_edge))
            # connection between child nodes 
            nn_v1 <- as_ids(neighbors(g, v1, mode = c( "all")))
            nn_old_nei <- as_ids(neighbors(g, old_nei, mode = c( "all")))
            random_son_v1 <- sample(1:length(nn_v1), 1)
            random_son_old_nei <- sample(1:length(nn_old_nei), 1)
            # a node cannot be chosen 2 times 
            while(V(g)$busy[as.integer(nn_v1[random_son_v1])] == TRUE) {
                random_son_v1 <- sample(1:length(nn_v1), 1)
            }
            while(V(g)$busy[as.integer(nn_old_nei[random_son_old_nei])] == TRUE) {
                random_son_old_nei <- sample(1:length(nn_old_nei), 1)
            }
            # adding the connection between child nodes 
            g <- add_edges(g, edges = c(nn_v1[random_son_v1], nn_old_nei[random_son_old_nei]))
            # set the busy = TRUE to avoid multiple connections 
            V(g)$busy[as.integer(nn_v1[random_son_v1])] <- TRUE
            V(g)$busy[as.integer(nn_old_nei[random_son_old_nei])] <- TRUE
          }
        }
      }
      V(g)$done[v1] <- TRUE
    } 
  }
  nn_list <- NULL
  # Saving some useful qualities 
  diameter <- append(diameter, length(get_diameter(g)))
  numberNodes <- append(numberNodes, vcount(g))
}
```

#### Plot of the produced network.

```{r}
plot(g, 
     vertex.label=NA, 
     vertex.color="green",
     vertex.size=2,
     edge.color="red",
     edge.width=1)
```

## Box-covering algorithm

Covering a network with the minimum possible number of boxes can reveal precious features of the network's structure. In our case, we seek to identify clues regarding self-similar or fractal features. This task can be accomplished using a box-covering algorithm. The ultimate aim of all box-covering algorithms is to identify the optimal solution, i.e., to find the minimum numeber of boxes for any given box size. A possible way consists of mapping the problem into the well-known greedy color algorithm, a widely adopted technique to obtain an approximate solution.

### Dual graph and greedy-color algorithm:

Eventually, we ended up in a configuration where the coloring in graph h produces a natural box covering in the original network g, in the sense that vertices of the same color will necessarily form a box since the distance between them must be less than lb. In the very end, the minimum number of boxes is equal to the minimum required number of colors in the dual network h.

```{r}
# Dual graph 
lb_max <- 5 # maximum value of the box radius 
N_box <- c() 
for (lb in 1:lb_max){
  # initializtion of the dual graph 
  h <- delete_edges(g, E(g))
  for (v1 in (1:vcount(h))){
    nn <- neighbors(graph = h, v = v1)
    for (v2 in (1:vcount(h))){
      shortest_path <- shortest_paths(g, from = v1, to = v2)$vpath[[1]]
      dist_ij <- length(shortest_path)
      # CHECK - if the distanze between nodes is smaller than the size of the box
      if((dist_ij > lb) && !(v2 %in% nn )) {
        h <- add_edges(h, edges = c(v1, v2))
      }
    }
  }
  # Box covering: greedy-color 
  # set a NULL(-1) color for each node 
  color <- rep(-1, vcount(h)) # list containing the color of each node 
  B <- lapply(1:vcount(h), function(i) neighborhood(h, order = 1, nodes = i))
  color[1] <- 0 
  palette <- 0:(vcount(h)-1) # list of all available colors
  #
  for (v in V(g)){
    v <- as.numeric(v)
    # the first node has to be skipped 
    if (v == 1){
      next
    }
    else{
      # Assign the smallest available color that has not yet been picked by its neighbors. 
      neig <- B[[v]][[1]][-1]
      col_neig <- color[neig] 
      minimum <- min(palette[!(palette %in% col_neig)])  
      color[v] <- minimum
    }
  }
  # 
  N_box <- append(N_box, max(color) + 1)
}
```

The number of boxes needed to cover the network is extracted as the maximum value in the color list (list containing the color of each node) plus 1, since the color starts from 0 and not 1.
